Copulas and their potential for ecology

### 2019-10-13 Paper.Rmd
added a reference

<!--Partial Person correlation-->
We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
mentioned in the Inrtoduction, Background, and methods sections and 
reviewed here; but under some circumstances it may also be appropriate to use 
techniques which parallel our nonparametric statistics but that use unranked data, 
$x_i$, $y_i$. As mentioned previously, @Genest2007 and others
recommend carrying out inferences about dependence structures (which was our
goal) using normalized ranks, stating explicitly "statistical inference concerning
dependence structures should always be based on ranks." The reason for this is
that measures of dependence which use unranked data conflate information about
the maginal distributions of variables with information about the dependence
between variables. To provide an ecological example, suppose the true population densities $p_{1,t}$
and $p_{2,t}$ of two species of fish in a lake in year $t$ are unknown,
but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
science. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be more
or less correlated through time than the true densities would be, if they were known,
if the function $f$ relating $p_{i,t}$ and $c_{i,t}$ is nonlinear and if Pearson
correlation is used. The function $f$ modifies the marginal distributions, which modifies
measures such as Pearson correlation that conflate dependence and marginal information.
Rank-based measures such as Spearman or Kendall correlation, or any of our copula approaches,
will be unaffected by nonlinear monotonic functions such as $f$. The same difficulty will pertain in any
case for which a measurement of an ecological quantity is a nonlinear index of the true quantity
of interest. Nevertheless, if one is certain that measurements are linearly related to 
the true quantities being measured, statistics based on unranked data may be more appropriate 
under some circumstances, in part because ecological mechanisms which produce relationships 
between variables are influenced not only by the ordinal informatio in ranks but also 
by the cardinal information in the unranked data. Judgements of what statistics
are appropriate depend, of course, on the purpose of the analysis; the goal of this 
study was to study dependence in isolation, so we used ranks. If one is interested
in tail associations using cardinal data, a partial Pearson correlation can developed
straightforwardly by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
by the $x$ and $y$, but still computing the sum in that equation over data points
with ranks, $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$. This would result in a different approach to
tail associations that would conflate dependence and marginal information but 
that may also be useful for understanding ecological relationships.

>>>>

<!--Partial Person correlation-->
We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
mentioned in the Introduction, Background, and methods sections and 
reviewed here; but under some circumstances it may also be appropriate to use 
techniques which parallel our nonparametric statistics but that use unranked data, 
$x_i$, $y_i$. As mentioned previously, @Genest2007 and others
recommend carrying out inferences about dependence structures (which was our
goal) using normalized ranks, stating explicitly "statistical inference concerning
dependence structures should always be based on ranks." The reason for this is
that measures of dependence which use unranked data conflate information about
the maginal distributions of variables with information about the dependence
between variables. For instance, suppose the true population densities $p_{1,t}$
and $p_{2,t}$ of two species of fish in a lake in year $t$ are unknown,
but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be more
or less correlated through time than the true densities would be, if they were known,
if the function $f$ relating $p_{i,t}$ and $c_{i,t}$ is nonlinear and if Pearson
correlation is used. The function $f$ modifies the marginal distributions, which modifies
measures such as Pearson correlation that conflate dependence and marginal information.
Rank-based measures such as Spearman or Kendall correlation, or any of our copula approaches,
will be unaffected by nonlinear monotonic functions such as $f$. The same difficulty will pertain in any
case for which a measurement of an ecological quantity is a nonlinear index of the true quantity
of interest. Nevertheless, if one is certain that measurements are linearly related to 
the true quantities being measured, statistics based on unranked data may be more appropriate 
under some circumstances, in part because ecological mechanisms which produce relationships 
between variables are influenced not only by the ordinal information in ranks but also 
by the cardinal information in the unranked data. Judgements of what statistics
are appropriate depend, of course, on the purpose of the analysis; the goal of this 
study was to study dependence in isolation, so we used ranks. If one is interested
in tail associations using cardinal data, a partial Pearson correlation can developed
straightforwardly by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
by the $x$ and $y$, but still computing the sum in that equation over data points
with ranks, $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$. This would result in a different approach to
tail associations that would conflate dependence and marginal information but 
that may also be useful for understanding some ecological relationships.
